The prime of their life - the halflife of acting talent

### 2017-08-27 01_actor_careers/05_report.Rmd
updates

---
title: "Films: Age & Gender"
output: 
html_document:
code_folding: hide
---

# Questions?

Let's say we had a couple of questions about actors and directors:

> Do actors and directors improve with age?

> Are female actors generally cast  than their male counterparts?

To answer these questions we're going to need some data.

@@

I started with imdb user **hagennelson's** list of [top 1000 actors](http://www.imdb.com/list/ls058011111/) and other imdb user **earthanime's** list of [top 250 directors](http://www.imdb.com/list/ls062419846/). Both lists span across generations, from the ancient (Ethel Barrymore and Alfred Hitchcock) to the modern (Chloë Grace Moretz and Christopher Nolan).

I then developed code (which I'll cover in a different post) to scrape data from the rotten tomatoes pages of the above individuals' films. The final dataset includes **12,258** films and **1,233** indviduals.

Altough we'll only use a few of these variables I brought in a few extras for some future investigations:

Variable | Description
-------- | -----------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seems like screen time)

@@

We'll load and take a peek at the film dataset:

@@

@@

Unfortunately, because the Rotten Tomatoes devs are a bunch of politically correct SJWs, gender isn't provided on their film bio pages so we'll have to be creative. The `gender` package in R allows you to infer a person's gender from their first name and DOB. The result is a probabilistic best guess of that individuals most likely gender given their first name and DOB.

@@

# Female Directors?

# How old?

>>>>

---
title: "The prime of their life - the halflife of acting talent"
output: 
html_document:
code_folding: hide
---

> insert an image

Much analysis has been completed regarding the performing peak of athletes and sportspeople. Statistically, swimmers peak at around 24, marathoner runners between 30-31, gymnasts at a mere 18, NBA players generally peak between 26-29, professional hockey players between 27-28, and chess players at 30-35.

Does the idea of a performing prime extend to artists and specifically actors and directors? There are obvious reasons to be skeptical. There's no major physical component to acting that could decay as the body does, and unlike chess you wouldn't think there's the need for an exreme state of concentration and cognition which might be unreachable in older age. There are other confounding factors that might prevent us from answering this question difinitively (which I'll discuss later) but let's dig into the data.

@@

To curate a data set for anlaysis I started with imdb user **hagennelson's** list of [top 1000 actors](http://www.imdb.com/list/ls058011111/) and imdb user **earthanime's** list of [top 250 directors](http://www.imdb.com/list/ls062419846/). Both lists span across generations, from the ancient (Ethel Barrymore and Alfred Hitchcock) to the modern (Chloë Grace Moretz and Christopher Nolan).

I then developed code (available on GitHub) to scrape data from the rotten tomatoes pages of the above individuals' films. The final dataset includes **12,258** films and **1,233** indviduals. Although we'll only use a few of these variables, I brought in a few extras for some future investigations:

Variable | Description
-------- | -----------
url | URL extension on rotten tomatoes website
year | Film production year
title | Film title
reviews | Count of formal critic reviews of film
tomatometer | Percentage of favourable reviews
av_rating | Average rating of film out of 10
audience | The non-critic plebian equivolent of the tomatometer
box_office | The box-office revenue
director | The film's director
actor_1:actor_6 | The 6 top actors according to rotten tomatoes internal ranking (seemingly ordered by screen time)

@@

First we'll load and take a peek at the film dataset:

@@

films %>% head()

@@

Unfortunately, because the Rotten Tomatoes development tea are a bunch of politically correct SJWs, gender isn't provided on their film bio pages so we'll have to be creative. The `gender` R package allows you to infer a person's gender from their first name and DOB. The result is a probabilistic estimate of that individuals most likely gender given their first name and DOB.

@@

# Where are the female directors?

To add a layer to this analysis I wanted group together people by gender and film production role (or actor vs director). So we think we'd be working with 4 groups, however, I was suprised to see that of our 250 top directors only XXX were female which is pretty appaling.

This got me thinking about the rest of the dataset. Given our **12,258** films dataset isn't limited to director credits in our top 250 (the film might just have a top 1000 actor appear in it) we can get a bit of a better idea of the broader gender split of directors. I've captured a lot of the key numbers in the following chart.

```{r}
```

# So
